---
title: "The Spark Foundation"
author: "Bipul Jha"
subtitle: Task 4 - Bussiness Analytics
output: pdf_document
---

```{r warning_off}
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(ggplot2)
```

## **1. The Data**

The data has been downloaded and it is in the **.csv** file. We create a variable called **store_data** and extract the contents of the dataset in that variable.
```{r}
store_data <- read.csv("SampleSuperstore.csv")
```

### 1.1 Checking for any missing data in the dataset:

We take the sum of the output of **is.na(store_data)**. There can be 2 possible output's: **TRUE** and **FALSE**, and we know that the value assigned to **TRUE** is 1 and that to **FALSE** is 0. So, taking a sum of the function will give us the number of **missing values** in the entire dataset.

```{r}
sum(is.na(store_data))
```

As we see that the output of the previous function is **zero**, so, there are **no missing values** in the entire dataset.

### 1.2 Looking at the data:

We look at the summary of the data, it will give us the **dimenssions**, **columns** and type of the **store_data** dataset.

```{r}
summary(store_data)
```

We look at the first 6 rows of the **store_data** datset now, to see how the data looks like.

```{r}
head(store_data)
```

Now we check the **dimenssions** of the data set

```{r}
dim(store_data)
```

We see that there are **9994 rows** and **13 columns** in the dataset.

## **2. Exploratory Analysis**

Now it is the time we start exploring the dataset. One of the key things that we need to check is the **profit** that was gained out of each state. First we need to aggregate the dataset and store the result in a new variable named as **profit_state**, and check it's dimenssions.

```{r}
profit_state <- aggregate(Profit~State, data = store_data, sum)
dim(profit_state)
```

Now we make the first plot and check the Profit earned in each state.

```{r}
barplot(profit_state$Profit, names.arg = profit_state$State, xlab = "States", ylab = "Profit", main = "State VS Profit")
```

As we see, all the state names aren't visible, so we will try to make a plot for **profit** in each **region** of the country.


```{r}
region_profit <- aggregate(Profit~Region, data = store_data, sum)
barplot(region_profit$Profit, names.arg = region_profit$Region, xlab = "Region", ylab = "Profit", main = "Region VS Profit")
```

This shows that the maximum profit was earned in the **west** region, and hence, making an investment in that region can be fairly successful.

Now we will try to make a comparision of the data for **Profit** on each **State** based on **Region**

```{r}
region_state_profit <- aggregate(Profit~Region+State, data = store_data, sum)
barplot(region_state_profit$Profit, names.arg = region_state_profit$State, col = region_state_profit$Region, main = "Profit of the States as per the Regions")
legend("topright", legend = unique(region_state_profit$Region), col = c("green", "blue", "red", "black"), pch =  19)
```

Making another graph just to demonstrate the comparision of **Region**, **State** and **Profit**

```{r}
qplot(Region, Profit, data = region_state_profit, color = State, main = "Profit based on Region and States")
```

We now make 2 new variables that store the profit and loss in each state. The variables are named as **profit** and **loss** respectively.

```{r}
loss <- subset(store_data, Profit <= 0)
profit <- subset(store_data, Profit > 0)
```

We now check for the **dimenssions** of both the variables, i**profit** and **loss**.

```{r}
dim(loss)
```

The dimenssion of **loss** came out to be 1936*13.

```{r}
dim(profit)
```

The dimenssion of **profit** came out to be 8058*13.

Now we make the plot for just the **profit** data of each **State**

```{r}
qplot(Profit, State, data = profit, alpha = 0.5, main = "Profit VS State")
```

So we see that higest single **profit amount** was seen in the **State of Indiana**. Now we go on to plot the the individual **loss** amount data as per each state.

```{r}
qplot(Profit, State, data = loss, alpha = 0.5, main = "Loss VS State", xlab = "Loss")
```

We see that the highest **individual loss** was incurred in the **State of Ohio**. But this dosent mean that the highest total loss was incurred in the same state. For that we aggregate the **profit** and **loss** data as per the states.

```{r}
profit_state <- aggregate(Profit~State+Region, data = profit, sum)
loss_state <- aggregate(Profit~State+Region, data = loss, sum)
```

Now we will make a plot for both **total profit** and **total loss** according to the states.

```{r}
qplot(Profit, State, data = profit_state, main = "Total Profit VS State", color = Region)
qplot(Profit, State, data = loss_state, main = "Total Loss VS State", xlab = "Loss", color = Region)
```

So we see that the **highest Profitable State** is **California** followed by **Ney York** and the state with the higest **loss** is **Texas**.

Now we make a more informative plot for the **Region** and the **Profit** data

```{r}
qplot(Profit, data = region_state_profit, fill = Region, main = "Profit VS Region")
```

So we see that most of the profit for the **Central** region is focussed around **Zero**

Now we check the **Profit** based on the **Category** and **Sub Category**

```{r}
profit_sub_cat <- aggregate(Profit ~ Category+Sub.Category, data = store_data, sum)
barplot(profit_sub_cat$Profit, names.arg = profit_sub_cat$Sub.Category, main = "Profit VS Sub Category", xlab = "Sub Category", ylab = "Total Profit")
qplot(Profit, Sub.Category, data = store_data, alpha = 0.5, color = Category, ylab = "Sub Category", main = "Profit VS Sub Category")
qplot(Profit, Sub.Category, data = profit_sub_cat, main = "Total Profit VS Sub Category", ylab = "Sub Ctaegory", xlab = "Total Profit", color = Category)
```

We see that **Copier** gives the higest profit and **Tables** give the highest discount.  

Now we will make graphs related to the **quantity** of the product sold based on the **Category**, which will show us what are the products that we should focus on for **manufacturing** purposes.  

First we will see how much **Quantity** is required in each **Segment**

```{r}
seg_quan <- aggregate(Quantity~Segment, data = store_data, sum)
barplot(seg_quan$Quantity, names.arg =  seg_quan$Segment, xlab = "Segment", ylab = "Quantity sold", main = "Quantity VS Segment")
```

As we see the largest **Quantity** requirement is by the **Consumer** segment.  

Now we focus on the category of the product and the **Quantity** of sale of each product.
```{r}
quan_sub_cat <- aggregate(Quantity ~ Category+Sub.Category, data = store_data, sum)
qplot(Quantity, Sub.Category, data = quan_sub_cat, color = Category, main = "Quantity VS Sub Category", ylab = "Sub Category")
```

Now we check for the **Quantity** of each **sub categpory** as per the **Segment**

```{r}
sub_seg_quan <- aggregate(Quantity ~ Sub.Category+Segment, data = store_data, sum)
qplot(Quantity, data = sub_seg_quan, fill = Sub.Category, facets = Segment~., main = "Quantity of the Segments")
```

Now we move on to checking the **profit** based on **Category**, **Sub Category** and **Segment**.  
First we move on to the **profit** provided by each **segment**

```{r}
seg_profit <- aggregate(Profit~Segment, data = store_data, sum)
barplot(seg_profit$Profit, names.arg = seg_profit$Segment, xlab = "Segments", ylab = "Total Profit", main = "Profit in Each Segment")
```

Now we move on to findind the **Profit** as per the **Category** and **Segment**

```{r}
qplot(Profit, Segment, data = store_data, color = Category, alpha = 0.5)
seg_cat_profit <- aggregate(Profit~Segment+Category, data = store_data, sum)
qplot(Profit, data = seg_cat_profit, fill = Category, facets = Segment~., main = "Profit according to Segments")
```


```{r}
ggplot(store_data, aes(log10(Profit), Segment, color = Category)) + geom_boxplot() + labs(title = "Profit VS Segment", x = "log of Profit")
```

Checking the **Shipping Mode** with the **Profit** can be a key aspect of the **Explorotory Analysis**.  

```{r}
ship_profit <- aggregate(Profit~Ship.Mode, data = store_data, sum)
barplot(ship_profit$Profit, names.arg = ship_profit$Ship.Mode, xlab = "Shipping Mode", ylab = "Total Profit", main = "Profit VS Shipping Mode")
```

We now look at the explore the **Shipping Mode** data with **Profit** based on the **Category** and **Region**.

```{r}
ggplot(data = store_data, aes(Ship.Mode, Profit, color = Category)) + geom_point(aes(alpha = 0.5)) + facet_wrap(~Region, nrow = 4, ncol = 1) + labs(title = "Shipping Mode and Profit according to the Category as per Region", x = "Shipping Mode")
```

We now move on to checking the **Profit** based on the **Category**, **Segment** and **Region**, and we do that by making a new variable, namely **profit_cat_seg_reg**.

```{r}
profit_cat_seg_reg <- aggregate(Profit ~ Region+Category+Segment, data = store_data, sum)
ggplot(profit_cat_seg_reg, aes(x = Region, y = Profit, color = Category)) + geom_point() + facet_wrap(~Segment, nrow = 3, ncol = 1) + labs(title = "Profit per Region as per Category and Segment")
```

#### We see that one of the highest profits in each section are:  
##### Consumer  
1. **West** - Higest profit by *Technology* followed by *Office Supplies* and *Furniture*  
2. **South** - Higest profit by *Technology* followed by *Office Supplies* and *Furniture*  
3. **East** - Higest profit by *Technology* followed by *Office Supplies* and *Furniture*  
4. **North** - Higest profit by *Technology* followed by *Office Supplies* and *Furniture*  

##### Corporate
1. **West** - Higest profit by *Office Supplies* followed by *Technology* and *Furniture*  
2. **South** - Higest profit by *Office Supplies* followed by *Technology* and *Furniture*  
3. **East** - Higest profit by *Office Supplies* followed by *Technology* and *Furniture*  
4. **North** - Higest profit by *Technology* followed by *Furniture* and *Office Supplies*

##### Home Office
1. **West** - Higest profit by *Office Supplies* followed by *Technology* and *Furniture*  
2. **South** - Higest profit by *Office Supplies* followed by *Furniture* and *Technology*  
3. **East** - Higest profit by *Technology* followed by *Office Supplies* and *Furniture*  
4. **North** - Higest profit by *Office Supplies* followed by *Technology* and *Furniture*  

## **3. Summary**
### Few key points that we notice based on the Profit:
1. **Profit** is maximum in the **Western Region** of the country.  
2. The maximum profits in all the **States** comes from the **State of California(West)** followed by **New York(East)**
3. The maximum profit was earned by selling the **copiers**.
4. The Quantity of product bought was most by the **Consumers segment** followed by **Corporate** and **Home Offices segments**
5. The products sold in the:
a. **Furnishing Category**: Maximum was **Furnishings**, Minimum was **Bookcases**  
b. **Office Supplies**: Maximum was **Binders**, Minimum was **Tables**  
c. **Technology**: Maximum was **Phones**, Minimum was **Copiers**
6. For all the **Segments**, **Technology is the most profitable category**, followed by **Office Supplies** and **Furniture**.
7. We see that the **Standard Shipping** is the most profitable **Shipping Mode** in all the **Categories**. But it is important to note that **Data** provided for making an analysis of the **Shipping Mode** depence on **Profit** is quite **insufficient**, because the **Shipping mode** depends on a lo*t of other external factors.

### Few key points that we notice based on the Loss:
1. **Profit** is minimum in the **Central Region** of the country.
2. The states of **Texas(Central)**, **Pensylvania(East)** and **Ohio(East)** incurred the highest amount of **loss.**
3. Highest **Losses** are incurred in selling the **Tables(Furniture)**, **Bookcases(Furniture)** and **Supplies(Office Supplies)** respectively. 
4. We see that the **Quantity** of **Copiers(Technology)** sold is the least, but still it was able to give the highest **profit**. This is a key point, and the copany should advertise the **Copier** more and produce it in ample amount, so that they can increase their **profit**.
5. We see that the profit for by the **Furniture Category** is least in all the **Segments**, so, we can use this analysis to make the **Furniture** only when ordered and as **Furniture** is a **depleting comodity**, it is advisable not to store it for a long time. The company can hence invest the money saved into making new **Technology**, so that **more profit can be made out of the same investment**.
